---
title: "Final Project - Freddie"
author: "Freddie Kiessling"
date: "2023-12-05"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# load libraries
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(dplyr)
library(tidyr)
library(corrplot)
```

```{r}
# load datasets
orig_train = read.csv("/Users/freddiekiessling/Downloads/train.csv")
orig_test = read.csv("/Users/freddiekiessling/Downloads/test.csv")
train = read.csv("/Users/freddiekiessling/Downloads/train.csv")
test = read.csv("/Users/freddiekiessling/Downloads/test.csv")
```

```{r}
print(colnames(train))
```





# Data Cleaning/Pre-processing

As one of our goals is determining most significant predictors (not necessarily determining predicting the sale price of the home), in the data cleaning section I will be removing/combining many of the minor variables in order to help simplify the data set.

```{r}
# removing id column as irrelevant
train = train[,-1]
```

```{r}
# counting missing values
missing =  sapply(train, function(x) {
  sum(is.na(x)) / length(x) * 100
})
missing_df = data.frame(Missing_Percentages = missing)
missing_df = missing_df %>%
  arrange(desc(Missing_Percentages))
missing_df = missing_df[missing_df$Missing_Percentages > 0, , drop = F]
missing_df # percentage of columns with NA values
```

### Dealing with Numerical Variables

```{r}
# Convert character ratings to numerical, used in many variables
rating_to_score <- function(rating) {
  if (is.na(rating)) {
    return(0)
  } else {
    scores <- c(Ex = 5, Gd = 4, TA = 3, Fa = 2, Po = 1)
    return(scores[rating])
  }
}
```

```{r}
# Variables leaving as is: MSSubClass, LotArea, OverallQual, OverallCond, TotalBsmtSF, BedroomAbvGr, KitchenAbvGr, TotRmsAbvGrd, Fireplaces, GarageCars, GarageArea
```

```{r}
# Pool
# Add variable to check for pool or not due to large amount of missing values
train$HasPool <- ifelse(train$PoolArea > 0 | !is.na(train$PoolQC), 1, 0)

# Removing others due to missing variables
train$PoolArea <- NULL
train$PoolQC <- NULL
``` 

```{r}
# LotFrontage
# Due to unknown meaning of NA, removing:
train$LotFrontage <- NULL
```

```{r}
# Date variables
# Not needed
train$MoSold <- NULL

# Age of house when sold
train$AgeAtSale <- train$YrSold - train$YearBuilt

# Age since last remodel when sold
train$YearsSinceRemodel <- train$YrSold - train$YearRemodAdd

# Removing due to some missing values
train$GarageYrBlt = NULL
```

```{r}
# Masonry Veneer Area
train$MasVnrArea[is.na(train$MasVnrArea)] <- 0
```

```{r}
# Basement Square Footage and Above Grade Square Footage

# Only keeping total square feet for basement
train$BsmtUnfSF = NULL
train$BsmtFinSF1 = NULL
train$BsmtFinSF2 =NULL

# Only keeping low quality square feet and total for above grade
train$X1stFlrSF = NULL
train$X2ndFlrSF = NULL
```

```{r}
# Basement Bathrooms, combining into one variable
train$BsmtBaths <- train$BsmtFullBath + train$BsmtHalfBath * 0.5
train$BsmtFullBath <- NULL
train$BsmtHalfBath <- NULL
```

```{r}
# Above grade bathrooms, also combining
train$TotalBaths <- train$FullBath + train$HalfBath * 0.5
train$FullBath <- NULL
train$HalfBath <- NULL
```

```{r}
# Outside square footage
# Combining all into one variable
train$TotalOutSideSF = train$WoodDeckSF + train$OpenPorchSF + train$EnclosedPorch + train$X3SsnPorch + train$ScreenPorch

train$WoodDeckSF = NULL
train$OpenPorchSF = NULL
train$EnclosedPorch = NULL
train$X3SsnPorch = NULL
train$ScreenPorch = NULL
```

### Dealing with Categorical Variables

```{r}
# MSZoning
price <- summarize(group_by(train, MSZoning),
          mean(SalePrice, na.rm=T))
zone_mapping <- c("FV" = 5, "RL" = 4, "RM" = 3, "RH" = 2, "C (all)" = 1)
train <- train %>%
  mutate(MSZoning = zone_mapping[MSZoning])
```

```{r}
summarize(group_by(train, Street),
          mean(SalePrice, na.rm=T))
# Street, changed to 1 if paved, 0 if not
train$Street <- ifelse(train$Street == "Pave", 1, 0)

summarize(group_by(train, Alley),
          mean(SalePrice, na.rm=T))
# Alley, 1 if alley access, 0 if not
train$Alley <- ifelse(is.na(train$Alley), 0, 1)
```

```{r}
# Lot Variables
summarize(group_by(train, LotConfig),
          mean(SalePrice, na.rm=T))
# 1 for sale price > $200000, 0 for < $200000
train$LotConfig = ifelse(train$LotConfig %in% c("CulDSac", "FR3"), 1, 0)

# Regular or not
train$LotShape <- ifelse(train$LotShape == "Reg", 1, 0)

# Flat or not
train$isFlat <- ifelse(train$LandContour == "Lvl", 1, 0)
train$LandContour <- NULL
```

```{r}
# Utilities, 1 for has everything, 0 for missing something
summarize(group_by(train, Utilities),
          mean(SalePrice, na.rm=T))
train$Utilities <- ifelse(train$Utilities == "AllPub", 1, 0)
train$Utilities <- NULL

# Electricity, excluding due to unknown variable meanings
train$Electrical = NULL
```

```{r}
# Land Slope
summarize(group_by(train, LandSlope),
          mean(SalePrice, na.rm=T))
# unexpected correlation, will keep all separate
slopes_mapping <- c("Gtl" = 1, "Mod" = 2, "Sev" = 3)
train <- train %>%
  mutate(LandSlope = slopes_mapping[LandSlope])
```

```{r}
# Neighborhoods

# Sorting by average sale price by neighborhood
average_prices <- train %>%
  group_by(Neighborhood) %>%
  summarize(AveragePrice = mean(SalePrice, na.rm = TRUE)) %>%
  ungroup() 

# Splitting into 3 classes, upper, middle, lower
quantiles <- quantile(average_prices$AveragePrice, probs = c(0.33, 0.66))

train <- train %>%
  left_join(average_prices, by = "Neighborhood")

train$NeighborhoodClass <- ifelse(
  train$AveragePrice <= quantiles[1], 1,  # Lower Priced Neighborhoods
  ifelse(train$AveragePrice > quantiles[1] & train$AveragePrice <= quantiles[2], 2, 3)  # Middle or Upper Priced Neighborhoods
)

train$AveragePrice <- NULL
train$Neighborhood <- NULL
```

```{r}
# Condition Proximity
summarize(group_by(train, Condition1),
          mean(SalePrice, na.rm=T))

train$NearPositiveFeature1 <- ifelse(train$Condition1 %in% c("PosA", "PosN"), 1, 0)
train$NearPositiveFeature2 <- ifelse(train$Condition1 %in% c("PosA", "PosN"), 1, 0)

train$NearRailroad1 <- ifelse(train$Condition1 %in% c("RRNn", "RRAn", "RRNe", "RRAe"), 1, 0)
train$NearRailroad2 <- ifelse(train$Condition1 %in% c("RRNn", "RRAn", "RRNe", "RRAe"), 1, 0)

train$Condition1 <- NULL
train$Condition2 <- NULL
```

```{r}
# Building Type
summarize(group_by(train, BldgType),
          mean(SalePrice, na.rm=T))

# Combining townhouse and spltting into 4 different binary variables
train$One_family_building <- ifelse(train$BldgType == "1Fam", 1, 0)
train$Two_family_conversion <- ifelse(train$BldgType == "2fmCon", 1, 0)
train$Duplex <- ifelse(train$BldgType == "Duplex", 1, 0)
train$Townhouse <- ifelse(train$BldgType %in% c("TwnhsE", "Twnhs"), 1, 0)

train$BldgType <- NULL
```

```{r}
# House Style
summarize(group_by(train, HouseStyle),
          mean(SalePrice, na.rm=T))

train$Split_style <- ifelse(train$HouseStyle %in% c("SLvl", "SFoyer"), 1, 0) 
train$Less_than_two_story <- ifelse(train$HouseStyle %in% c("1Story", "1.5Fin", "1.5Unf"), 1, 0)
train$Two_story_plus <- ifelse(train$HouseStyle %in% c("2Story", "2.5Fin", "2.5Unf"), 1, 0)

train$HouseStyle <- NULL
```



```{r}
# Roof + Exterior + Masonry Veneer + Foundation
summarize(group_by(train, RoofStyle),
          mean(SalePrice, na.rm=T))
# I'm going to leave these out of the data set as I don't know what exactly all these are and can't classify as binary
train$RoofStyle <- NULL

summarize(group_by(train, RoofMatl),
          mean(SalePrice, na.rm=T))

train$RoofMatl <- NULL

# Same with Exterior + Foundation
train$Exterior1st <- NULL
train$Exterior2nd <- NULL
train$Foundation <- NULL

# Changing exterior condition and quality to numeric
train$ExterQual <- sapply(train$ExterQual, rating_to_score)
train$ExterCond <- sapply(train$ExterCond, rating_to_score)

summarize(group_by(train, MasVnrType),
          mean(SalePrice, na.rm=T))

train$hasMasVnr <- ifelse(is.na(train$MasVnrType), 0, 1)

train$MasVnrType <- NULL

```

```{r}
# Basement
train$BsmtQualScore <- sapply(train$BsmtQual, rating_to_score)
train$BsmtCondScore <- sapply(train$BsmtCond, rating_to_score)

train$BsmtQual <- NULL
train$BsmtCond <- NULL

# Exposure
train$BsmtExposure[train$BsmtExposure == "Gd"] <- 3
train$BsmtExposure[train$BsmtExposure == "Av"] <- 2
train$BsmtExposure[train$BsmtExposure == "Mn"] <- 1
train$BsmtExposure[train$BsmtExposure == "No"] <- 0
train$BsmtExposure[is.na(train$BsmtExposure)] <- 0

# Finished Types
train$BsmtFinType1[train$BsmtFinType1 %in% c("GLQ", "ALQ", "Rec")] <- 1 
train$BsmtFinType1[!train$BsmtFinType1 %in% c("GLQ", "ALQ", "Rec")] <- 0

train$BsmtFinType2[train$BsmtFinType2 %in% c("GLQ", "ALQ", "Rec")] <- 1 
train$BsmtFinType2[!train$BsmtFinType2 %in% c("GLQ", "ALQ", "Rec")] <- 0
```

```{r}
# Heating + CentralAir
# Leaving out heating types as unsure of quality
train$Heating <- NULL

# Changing quality to numeric
train$HeatingQC <- sapply(train$HeatingQC, rating_to_score)
train$HeatingQC <- NULL

# Converting to binary
train$CentralAir <- ifelse(train$CentralAir == "Y", 1, 0)
```

```{r}
# Kitchen 
# Changing quality to numeric
train$KitchenQual <- sapply(train$KitchenQual, rating_to_score)
```

```{r}
# Functionality
# Due to unclear guidelines, leaving this out
train$Function <- NULL
```

```{r}
# Fireplaces
# Changing quality to numeric
train$FireplaceQu <- sapply(train$FireplaceQu, rating_to_score)
```

```{r}
# Garage
train$GarageQual <- sapply(train$GarageQual, rating_to_score)
train$GarageCond <- sapply(train$GarageCond, rating_to_score)
train$HasGarage <- ifelse(is.na(train$GarageType), 0, 1)  

train$GarageType <- NULL
train$GarageFinish <- NULL
```

```{r}
# Paved driveway
train$PavedDrive[train$PavedDrive == "Y"] <- 1
train$PavedDrive[!train$PavedDrive != "Y"] <- 0
train$PavedDrive[is.na(train$PavedDrive)] <- 0
```

```{r}
# Fence 
train$Fence[train$Fence %in% c("GdPrv", "MnPrv", "GdWo","MnWw")] <- 1
train$Fence[is.na(train$Fence)] <- 0
```

```{r}
# MiscFeature and MiscVal variables
# Average value of miscellaneous feature
mean(train$MiscVal[train$MiscVal != 0])

# Due to the value being relatively modest compared to the average final sale price of the home, we will exclude it
train$MiscFeature <- NULL
train$MiscVal <- NULL
```

```{r}
# Sale Type + Sale Condition
# Unsure of quality of each type, excluding
train$SaleType = NULL
train$SaleCondition = NULL
```

Done with pre-processing

```{r}
# Double checking for missing values
missing =  sapply(train, function(x) {
  sum(is.na(x)) / length(x) * 100
})
missing_df = data.frame(Missing_Percentages = missing)
missing_df = missing_df %>%
  arrange(desc(Missing_Percentages))
missing_df = missing_df[missing_df$Missing_Percentages > 0, , drop = F]
missing_df # percentage of columns with NA values
```

```{r}
# Correlation Matrix for numeric variables
numeric_vars <- sapply(train, is.numeric)
cor_matrix <- cor(train[, numeric_vars], use = "pairwise.complete.obs")

corrplot(cor_matrix, method = "square",
         tl.cex = 0.5, 
         tl.srt = 90,  
         addrect = 5)
```

```{r}
# Correlation Matrix using all original data for reference (delete later)
numeric_vars <- sapply(orig_train, is.numeric)
cor_matrix <- cor(orig_train[, numeric_vars], use = "pairwise.complete.obs")

corrplot(cor_matrix, method = "square",
         tl.cex = 0.5, 
         tl.srt = 45,  
         addrect = 5)
```

As we made many changes to the training data all the steps must also be applied to the test data:

```{r}
# removing id column as irrelevant
test = test[,-1]
```

Dealing with Numerical Variables

```{r}
# Variables leaving as is: MSSubClass, LotArea, OverallQual, OverallCond, TotalBsmtSF, BedroomAbvGr, KitchenAbvGr, TotRmsAbvGrd, Fireplaces, GarageCars, GarageArea
```

```{r}
# Pool
# Add variable to check for pool or not due to large amount of missing values
test$HasPool <- ifelse(test$PoolArea > 0 | !is.na(test$PoolQC), 1, 0)

# Removing others due to missing variables
test$PoolArea <- NULL
test$PoolQC <- NULL
``` 

```{r}
# LotFrontage
# Due to unknown meaning of NA, removing:
test$LotFrontage <- NULL
```

```{r}
# Date variables
# Not needed
test$MoSold <- NULL

# Age of house when sold
test$AgeAtSale <- test$YrSold - test$YearBuilt

# Age since last remodel when sold
test$YearsSinceRemodel <- test$YrSold - test$YearRemodAdd

# Removing due to some missing values
test$GarageYrBlt = NULL
```

```{r}
# Masonry Veneer Area
test$MasVnrArea[is.na(test$MasVnrArea)] <- 0
```

```{r}
# Basement Square Footage and Above Grade Square Footage

# Only keeping total square feet for basement
test$BsmtUnfSF = NULL
test$BsmtFinSF1 = NULL
test$BsmtFinSF2 =NULL

# Only keeping low quality square feet and total for above grade
test$X1stFlrSF = NULL
test$X2ndFlrSF = NULL
```

```{r}
# Basement Bathrooms, combining into one variable
test$BsmtBaths <- test$BsmtFullBath + test$BsmtHalfBath * 0.5
test$BsmtFullBath <- NULL
test$BsmtHalfBath <- NULL
```

```{r}
# Above grade bathrooms, also combining
test$TotalBaths <- test$FullBath + test$HalfBath * 0.5
test$FullBath <- NULL
test$HalfBath <- NULL
```

```{r}
# Outside square footage
# Combining all into one variable
test$TotalOutSideSF = test$WoodDeckSF + test$OpenPorchSF + test$EnclosedPorch + test$X3SsnPorch + test$ScreenPorch

test$WoodDeckSF = NULL
test$OpenPorchSF = NULL
test$EnclosedPorch = NULL
test$X3SsnPorch = NULL
test$ScreenPorch = NULL
```

Dealing with Categorical Variables
```{r}
# MSZoning
zone_mapping <- c("FV" = 5, "RL" = 4, "RM" = 3, "RH" = 2, "C (all)" = 1)
test <- test %>%
  mutate(MSZoning = zone_mapping[MSZoning])
```

```{r}
# Street, changed to 1 if paved, 0 if not
test$Street <- ifelse(test$Street == "Pave", 1, 0)

# Alley, 1 if alley access, 0 if not
test$Alley <- ifelse(is.na(test$Alley), 0, 1)
```

```{r}
# Lot Variables
# 1 for sale price > $200000, 0 for < $200000
test$LotConfig = ifelse(test$LotConfig %in% c("CulDSac", "FR3"), 1, 0)

# Regular or not
test$LotShape <- ifelse(test$LotShape == "Reg", 1, 0)

# Flat or not
test$isFlat <- ifelse(test$LandContour == "Lvl", 1, 0)
test$LandContour <- NULL
```

```{r}
# Utilities, 1 for has everything, 0 for missing something
test$Utilities <- ifelse(test$Utilities == "AllPub", 1, 0)
test$Utilities <- NULL

# Electricity, excluding due to unknown variable meanings
test$Electrical = NULL
```

```{r}
# Land Slope
# unexpected correlation, will keep all separate
slopes_mapping <- c("Gtl" = 1, "Mod" = 2, "Sev" = 3)
test <- test %>%
  mutate(LandSlope = slopes_mapping[LandSlope])
```

```{r}
# Neighborhoods
test$NeighborhoodClass <- ifelse(test$Neighborhood %in% c("NoRidge", "NridgHt", "StoneBr", "Timber", "Veenker"), 3, 
                                 ifelse(test$Neighborhood %in% c("ClearCr", "CollgCr", "Crawfor", "Somerst", "Gilbert", "NWAmes"), 2, 
                                        1))

test$Neighborhood <- NULL
```

```{r}
# Condition Proximity
test$NearPositiveFeature1 <- ifelse(test$Condition1 %in% c("PosA", "PosN"), 1, 0)
test$NearPositiveFeature2 <- ifelse(test$Condition1 %in% c("PosA", "PosN"), 1, 0)

test$NearRailroad1 <- ifelse(test$Condition1 %in% c("RRNn", "RRAn", "RRNe", "RRAe"), 1, 0)
test$NearRailroad2 <- ifelse(test$Condition1 %in% c("RRNn", "RRAn", "RRNe", "RRAe"), 1, 0)

test$Condition1 <- NULL
test$Condition2 <- NULL
```

```{r}
# Building Type
# Combining townhouse and spltting into 4 different binary variables
test$One_family_building <- ifelse(test$BldgType == "1Fam", 1, 0)
test$Two_family_conversion <- ifelse(test$BldgType == "2fmCon", 1, 0)
test$Duplex <- ifelse(test$BldgType == "Duplex", 1, 0)
test$Townhouse <- ifelse(test$BldgType %in% c("TwnhsE", "Twnhs"), 1, 0)

test$BldgType <- NULL
```

```{r}
# House Style
test$Split_style <- ifelse(test$HouseStyle %in% c("SLvl", "SFoyer"), 1, 0) 
test$Less_than_two_story <- ifelse(test$HouseStyle %in% c("1Story", "1.5Fin", "1.5Unf"), 1, 0)
test$Two_story_plus <- ifelse(test$HouseStyle %in% c("2Story", "2.5Fin", "2.5Unf"), 1, 0)

test$HouseStyle <- NULL
```

```{r}
# Roof + Exterior + Masonry Veneer + Foundation
# I'm going to leave these out of the data set as I don't know what exactly all these are and can't classify as binary
test$RoofStyle <- NULL
test$RoofMatl <- NULL

# Same with Exterior + Foundation
test$Exterior1st <- NULL
test$Exterior2nd <- NULL
test$Foundation <- NULL

# Changing exterior condition and quality to numeric
test$ExterQual <- sapply(test$ExterQual, rating_to_score)
test$ExterCond <- sapply(test$ExterCond, rating_to_score)

test$hasMasVnr <- ifelse(is.na(test$MasVnrType), 0, 1)

test$MasVnrType <- NULL

```

```{r}
# Basement
test$BsmtQualScore <- sapply(test$BsmtQual, rating_to_score)
test$BsmtCondScore <- sapply(test$BsmtCond, rating_to_score)

test$BsmtQual <- NULL
test$BsmtCond <- NULL

# Exposure
test$BsmtExposure[test$BsmtExposure == "Gd"] <- 3
test$BsmtExposure[test$BsmtExposure == "Av"] <- 2
test$BsmtExposure[test$BsmtExposure == "Mn"] <- 1
test$BsmtExposure[test$BsmtExposure == "No"] <- 0
test$BsmtExposure[is.na(test$BsmtExposure)] <- 0

# Finished Types
test$BsmtFinType1[test$BsmtFinType1 %in% c("GLQ", "ALQ", "Rec")] <- 1 
test$BsmtFinType1[!test$BsmtFinType1 %in% c("GLQ", "ALQ", "Rec")] <- 0

test$BsmtFinType2[test$BsmtFinType2 %in% c("GLQ", "ALQ", "Rec")] <- 1 
test$BsmtFinType2[!test$BsmtFinType2 %in% c("GLQ", "ALQ", "Rec")] <- 0
```

```{r}
# Heating + CentralAir
# Leaving out heating types as unsure of quality
test$Heating <- NULL

# Changing quality to numeric
test$HeatingQC <- sapply(test$HeatingQC, rating_to_score)
test$HeatingQC <- NULL

# Converting to binary
test$CentralAir <- ifelse(test$CentralAir == "Y", 1, 0)
```

```{r}
# Kitchen 
# Changing quality to numeric
test$KitchenQual <- sapply(test$KitchenQual, rating_to_score)
```

```{r}
# Functionality
# Due to unclear guidelines, leaving this out
test$Function <- NULL
```

```{r}
# Fireplaces
# Changing quality to numeric
test$FireplaceQu <- sapply(test$FireplaceQu, rating_to_score)
```

```{r}
# Garage
test$GarageQual <- sapply(test$GarageQual, rating_to_score)
test$GarageCond <- sapply(test$GarageCond, rating_to_score)
test$HasGarage <- ifelse(is.na(test$GarageType), 0, 1)  

test$GarageType <- NULL
test$GarageFinish <- NULL
```

```{r}
# Paved driveway
test$PavedDrive[test$PavedDrive == "Y"] <- 1
test$PavedDrive[!test$PavedDrive != "Y"] <- 0
test$PavedDrive[is.na(test$PavedDrive)] <- 0
```

```{r}
# Fence 
test$Fence[test$Fence %in% c("GdPrv", "MnPrv", "GdWo","MnWw")] <- 1
test$Fence[is.na(test$Fence)] <- 0
```

```{r}
# MiscFeature and MiscVal variables

# Due to the value being relatively modest compared to the average final sale price of the home, we will exclude it
test$MiscFeature <- NULL
test$MiscVal <- NULL
```

```{r}
# Sale Type + Sale Condition
# Unsure of quality of each type, excluding
test$SaleType = NULL
test$SaleCondition = NULL
```


Hubert's code starts here:





```{r}
# visualizations

# 5 individual plots based on columns w/ highest correlation

boxplot1 <- ggplot(train, aes(x = as.factor(OverallQual), y = SalePrice)) + 
            geom_boxplot() + 
            ggtitle("Boxplot - Overall Quality vs Sale Price") +
            xlab("Overall Quality") +
            ylab("Sale Price")
plot2 <- ggplot(train, aes(x = GrLivArea, y = SalePrice)) + geom_point() + ggtitle("Plot 2")
boxplot3 <- ggplot(train, aes(x = as.factor(GarageCars), y = SalePrice)) + 
            geom_boxplot() + 
            ggtitle("Boxplot - Garage Cars vs Sale Price") +
            xlab("Garage Cars") +
            ylab("Sale Price")

plot4 <- ggplot(train, aes(x = GarageArea, y = SalePrice)) + geom_point() + ggtitle("Plot 4")
plot5 <- ggplot(train, aes(x = TotalBsmtSF, y = SalePrice)) + geom_point() + ggtitle("Plot 5")





plot_list = list(boxplot1, plot2, boxplot3, plot4, plot5)


for (plot in plot_list) {
    print(plot)
}
```

12 Highest predictors
```{r}
correlations <- cor(train[sapply(train, is.numeric)], train$SalePrice)
sorted_correlations <- sort(abs(correlations[,1]), decreasing = TRUE)
top_12_predictors <- names(sorted_correlations)[2:13]

print(top_12_predictors)
sorted_correlations[1:13]

```

We chose to apply a Forward Selection algorithm to find the best multiple linear regression model based on the highest correlation predictors to Sale Price (Y). We could have also used Backward elimination or stepwise, but we decided it would be too computationally intensive due to our large number of predictors. Thus, we started with a base model, and applied a forward selection to add predictors. 

```{r}
base_model <- lm(SalePrice ~ 1, data = train)
library(MASS)
best_model <- stepAIC(base_model, scope = list(lower = base_model, upper = ~ OverallQual + GrLivArea + ExterQual + KitchenQual + GarageCars + GarageArea + TotalBsmtSF), direction = "forward")
```


K-fold CV (10 Folds): 
```{r}
library(caret)
fitControl <- trainControl(method = "repeatedcv", number = 10, repeats = 3)


model1_formula <- SalePrice ~ OverallQual + GrLivArea + KitchenQual + GarageCars + TotalBsmtSF + ExterQual
model2_formula <- SalePrice ~ OverallQual + GrLivArea + KitchenQual + GarageCars + TotalBsmtSF
model3_formula <- SalePrice ~ OverallQual + GrLivArea + KitchenQual + GarageCars


set.seed(123)  # For reproducibility
model1_cv <- train(model1_formula, data = train, method = "lm", trControl = fitControl)

set.seed(123)
model2_cv <- train(model2_formula, data = train, method = "lm", trControl = fitControl)

set.seed(123)
model3_cv <- train(model3_formula, data = train, method = "lm", trControl = fitControl)

model1_cv$results
model2_cv$results
model3_cv$results

```



Diagnostics:


```{r}
model1 <- lm(model1_formula, data = train)

# Residuals vs Fitted Values
plot(model1$fitted.values, residuals(model1), 
     xlab = "Fitted Values", 
     ylab = "Residuals", 
     main = "Residuals vs Fitted Values")
abline(h = 0, col = "red")

# Normal Q-Q Plot
qqnorm(residuals(model1))
qqline(residuals(model1), col = "red")


# Shapiro-Wilk Test
shapiro.test(residuals(model1))

# install.packages("lmtest")
library(lmtest)

# Breusch-Pagan Test
bptest(model1)

# install.packages("lmtest")
library(lmtest)

# Breusch-Pagan Test
bptest(model1)

# install.packages("car")
library(car)

# Variance Inflation Factor
vif(model1)

# install.packages("car")
library(car)

# Plot for Outliers and Leverage
influencePlot(model1, id.method = "identify", main = "Influence Plot", sub = "Circle size is proportional to Cook's distance")


```




```{r}
print(colnames(train))

```


Logistic Regression: utilizing backward elimination.


```{r}

```



```{r}

# Categorizing Sale Price into High and Low based on the median
median_price <- median(train$SalePrice)
train$PriceCategory <- ifelse(train$SalePrice > median_price, "High", "Low")
train$PriceCategoryBinary <- ifelse(train$PriceCategory == "High", 1, 0)

# Assuming you have already loaded the MASS package for stepAIC
library(MASS)

# Fit the full logistic regression model with all the specified predictors
full_log_model <- glm(PriceCategoryBinary ~ OverallQual + GrLivArea + ExterQual + KitchenQual + GarageCars + GarageArea + TotalBsmtSF + Neighborhood + TotalBaths + BsmtQualScore + TotRmsAbvGrd + AgeAtSale, data = train, family = binomial)

# Apply backward elimination to find the most optimal model
optimal_log_model <- stepAIC(full_log_model, direction = "backward")

# View the summary of the optimal model
summary(optimal_log_model)


```



